{\setstretch{1.5}

I would like to express my sincere thanks and gratitude to my principal supervisor Professor Mahen Mahendran, for his guidance, support and encouragement throughout my research. I am also grateful to my associate supervisor Dr Anthony Ariyanayagam for not only guiding me throughout this research, but also for the help and support both in professional and personal levels. The support from both my supervisors were invaluable without which completing this thesis would have been a nightmare.

I am grateful to the Queensland University of Technology (QUT) for supporting me with the QUTPRA award for my research. I also wish to express my sincere thanks to the Australian Research Council (ARC), National Association of Steel Framed Housing (NASH) and Stoddart group for providing the required materialistic support to conduct this research. 

Many thanks to Barry Hume and Glenn Atlee and all the other supporting staffs at QUT Banyo Pilot Plant Precinct for providing their assistance in conducting the experimental study for this research. My heartiest thanks to my fellow researchers Edward, Jashnav, Nilakshi, Janarthanan, Myuran, Thananjayan, Sayilacksha, Yomal, Imran, Rusthi for their help and support during this research at professional and personal levels.My Sincere thanks to all the undergraduate students who assisted me with the experimental study.

Finally, I would like to express my gratitude to my parents for their support, encouragement and blessings and my family members for their extended support. I am grateful to my wife Niveditha, the woman behind my success for her unconditional love, care and support.

}